---
title: "Homework 4"
author: "Nishant Arora"
header-includes:
   - \usepackage{bbm}
   - \usepackage{dplyr}
output:
    html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=TRUE, echo=FALSE}
library(dplyr, quietly=TRUE, warn.conflicts=FALSE)
```
## Part 1: Compare to known click rate (pA = 0.5)
## 1a:
Derive an expression for Var(X) [EX has been done as an example]

$$Var(\bar{X}) = Var(\frac{1}{n}\sum _{i=1}^n\:x_i)$$
$$=\frac{1}{n^2}\sum \:_{i=1}^n\:Var\left(x_i\right)$$
$$=n\:\cdot \:p_A\left(1-p_A\right)\cdot \frac{1}{n^2}$$
$$=\frac{p_A\left(1-p_A\right)}{n}$$

$$\therefore Var(\bar{X}) = \frac{p_A\left(1-p_A\right)}{n}$$

## 1b:
Based on your derivation, compute values for EX and Var(X) based on pA = 0.5 and n = 50.

```{r eval=TRUE}
pa <- 0.5
n <- 50

exp_val_samp_mean <- pa
var_samp_mean <- (pa * (1 - pa)) / n

exp_val_samp_mean
var_samp_mean
```

## 1c:
Using the result above, you can now use the CLT by approximating the distribution of X as N(EX, sqrt(Var(X))) Based on this approximation, compute p(X > pˆB). Use the R function pnorm to compute this.

```{r eval=TRUE}
pb <- 0.6
n <- 50

exp_val_samp_mean_b <- pb
var_samp_mean_b <- (pb * (1 - pb)) / n

pnorm(0.5, mean=0.6, sd=sqrt(var_samp_mean), lower.tail=FALSE)
```

## 1d:
Should you reject the null hypothesis pB ≤ pA? Why?

Note: if p(X > val) > α then you reject the null hypothesis.

From pnorm in 1c we get 0.9213504 which is greater than the rejection level of 0.05. Therefore we reject the null hypothesis.

## 1e:
What if you had observed the same pˆB = 0.6 but with n = 100 samples. Should you reject the null hypothesis in this case? Why?

```{r, eval=TRUE}
pb <- 0.6
n <- 100

exp_val_samp_mean_1c <- pb
var_samp_mean_1c <- (pb * (1 - pb)) / n

pnorm(0.5, mean=0.6, sd=sqrt(var_samp_mean_1c), lower.tail=FALSE)
```

We get a value of 0.9793866 which is greater than the rejection level of 0.05. Therefore we rejet the null hypothesis.

## 1f:
What is the smallest value pˆB you should reject the null hypothesis with n = 100. Use the qnorm function for this. Denote this smallest value as qB.