#### Nishant Arora
#### CMSC 320
#### HW 1  


```{r knitr_setup, echo=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#### Data types


_1) Provide a URL to the dataset._

Dataset source: https://catalog.data.gov/dataset/most-popular-baby-names-by-sex-and-mothers-ethnic-group-new-york-city-8c742



_2) Explain why you chose this dataset._ 

I found the data set to be interesting since I visit the city often and was curious. The dataset is clear and concise so I can create a plot easily. The file size also was quite reasonable and not extremely large.


_3) What are the entities in this dataset? How many are there?_

The entitites are babies born between 2011 and 2014 in NYC. There are 22,035 entities in the dataset. 

_4) How many attributes are there in this dataset?_

There are 6 attributes: Year of Birth, Gender, Ethnicity, Child's First Name, Count, Rank.


_5) What is the datatype of each attribute (categorical -ordered or unordered-, numeric -discrete or continuous-, datetime, geolocation, other)? Write a short sentence stating how you determined the type of each attribute. Do this for at least 5 attributes, if your dataset contains more than 10 attributes, choose 10 of them to describe._

| Num |         Name         |          Type         |                            Description                             |
|-----|----------------------|-----------------------|--------------------------------------------------------------------|
|  1  |    `Year of Birth`   |    numeric-discrete   |             The year must be 2011, 2012, 2013, or 2014             |
|  2  |       `Gender`       | categorical-unordered |          Non-numeric, must be from the (unordered) set M, F        |
|  3  |      `Ethnicity`     | categorical-unordered | Non-numeric, must be from a finites set of (unordered) ethnicities |
|  4  | `Child's First Name` | categorical-unordered |                  Name is being used to build models                |
|  5  |       `Count`        |    numeric-discrete   |                    Must be an exact whole number                   |
|  6  |        `Rank`        |    numeric-discrete   |                    Must be an exact whole number                   |


_6) Write R code that loads the dataset using function `read_csv`. Were you able to load the data successfully? If no, why not?_

```{r load_data, message=FALSE}
library(tidyverse)

name_tab <- read_csv("/Users/nishant/Desktop/data_science/Most_Popular_Baby_Names_by_Sex_and_Mother_s_Ethnic_Group__New_York_City.csv")
name_tab %>% slice(1:10)
```

#### Wrangling

1) My dataset contains duplicate values, so my pipeline removes duplicate names with the same year, gender and ethnicity, it also displays the top 5 most popular names for Hispanic children born in NYC in 2013.


```{r pipeline}
top_5_tab <- name_tab %>%
  filter(`Year of Birth` == 2013) %>%
  select(Gender, Ethnicity, `Child's First Name`, Count) %>%
  unique() %>%
  group_by(Ethnicity="HISPANIC") %>%
  arrange(desc(Count)) %>%
  slice(1:5)
top_5_tab
```

#### Plotting

1) This plot shows the 5 most popular Hispanic baby names in NYC in 2013.

```{r plot}
top_5_tab %>%
  ggplot(aes(x=`Child's First Name`, y=Count)) + geom_bar(stat="identity") 
top_5_tab
```