```{r knitr_setup, echo=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

## Data types

_1) Provide a URL to the dataset._

https://catalog.data.gov/dataset/most-popular-baby-names-by-sex-and-mothers-ethnic-group-new-york-city-8c742

_2) Explain why you chose this dataset._ 

I found the data set to be interesting since I visit the city often and was curious.  The dataset is clear and concise so I can create a plot easily. The file size also was quite reasonable and not extremely large.

_3) What are the entities in this dataset? How many are there?_


_4) How many attributes are there in this dataset?_


_5) What is the datatype of each attribute (categorical -ordered or unordered-, numeric -discrete or continuous-, datetime, geolocation, other)? Write a short sentence stating how you determined the type of each attribute. Do this for at least 5 attributes, if your dataset contains more than 10 attributes, choose 10 of them to describe._

| Num | Name | Type | Description |
|-----|------|------|-------------|
| 1 | <name> | <type> | <descr> |

_6) Write R code that loads the dataset using function `read_csv`. Were you able to load the data successfully? If no, why not?_

```{r load_data, message=FALSE}
library(tidyverse)
# loading code
name_tab <- read_csv("Most_Popular_Baby_Names_by_Sex_and_Mother_s_Ethnic_Group__New_York_City.csv")
name_tab
```

## Wrangling

1) My pipeline... 

```{r pipeline}
# pipeline goes here
```

## Plotting

1) This plot... 

```{r plot}
# plot goes here
```